### 2016-05-08 writeup/manuscript/manuscript.tex
Pub-ready tables, more analysis of nms30

% PREAMBLE ====
\documentclass[preprint,3p,twocolumn]{elsarticle} % TODO: What model?

% Use times font
\usepackage{mathptmx}
\usepackage{lmodern}

% Math
\usepackage{amsmath}

% Tables, Figures
\usepackage{booktabs}
\usepackage{threeparttable}
% For line-breaking cells
\newcommand{\spc}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\usepackage{float}
\usepackage{caption}
\setlength{\tabcolsep}{10pt}

% Graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png}
\graphicspath{{../../figures/}}

% References
\usepackage{numcompress}
\biboptions{super}
\bibliographystyle{model3-num-names}

% DOCUMENT ====
\begin{document}
% FRONTMATTER ====
\begin{frontmatter}
\title{Parkinson's disease cluster analysis}
\author[c3]{Pablo Martinez-Martin}
\ead{pmartinez@isciii.es}
\author[bc]{Jesse Mu}
\ead{muj@bc.edu}
\author[cig]{Concha Bielza}
\ead{mcbielza@fi.upm.es}
\author[cig]{Pedro Larra\~naga}
\ead{plarranaga@fi.upm.es}
\address[c3]{Area of Applied Epidemiology, National Centre of Epidemiology and CIBERNED, Carlos III Institute of Health, Madrid, Spain}
\address[bc]{Department of Computer Science, Boston College, Chestnut Hill, Massachusetts, USA}
\address[cig]{Computational Intelligence Group, Polytechnic University of Madrid, Madrid, Spain}

\begin{abstract}
\emph{Background} \quad This is the background. \\
\emph{Methods} \quad  These are methods. \\
\emph{Findings} \quad These are findings. \\
\emph{Interpretation} \quad This is interpretation. \\
\emph{Funding} \quad This is funding.
\end{abstract}
\journal{The Lancet Neurology}
\end{frontmatter}

% BODY ====

\section{Introduction}

\section{Methods}

\subsection{Statistical analysis}

Out of the 951 patients in the study, we used listwise deletion to exclude 47 patients due to
missing measurements in the variables of interest, resulting in 904 remaining patients. All
symptoms were scaled to $z$-scores ($\mu = 0$, $\sigma = 1$) before clustering, and unscaled
afterwards. Analysis was conducted in R 3.2.4 (www.r-project.org).

\subsubsection{Cluster analysis}

$k$-means was used for cluster analysis. We performed two analyses on the
patients in the dataset: the first on the nine aggregate nonmotor symptom
domains and the four motor symptoms, and the second on the 30 individual
nonmotor symptoms collected with the four motor symptoms. Complete-linkage
hierarchical clustering on the 30 nonmotor symptoms was also performed to observe the
categorization of symptoms.

\subsubsection{Determining $k$}

Various formal measures were used to determine the optimal number of
clusters for the dataset. First, the optimal $k$ according to the Gap Statistic and the
1-standard-error method\cite{tibshirani01gap} was $k = 4$ (Figure~\ref{fig:gap}). Second, although
the globally-optimal clusters reported from stability measures of the clValid R package
\cite{brock11clvalid} stability are 2, 6, and 8 (Figure~\ref{fig:stability}), there appeared to be
consensus that $k = 4$ is a locally-optimal number of clusters which supported the gap statistic
result.

Other cluster validation methods (SSE scree plot, minimum average silhouette width) generally
suggested 2 clusters. As a consequence, $k = 2, 3, 4$ was tried, where $k = 2, 3$ simply divided
the data uninformatively into groups with varying levels of overall PD severity. Thus $k = 4$ was
selected to offer a good blend of model fit, informativeness, and parsimony.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.pdf}
  \caption{Plot of gap statistic versus number of clusters with $k$-means on 100 bootstrapped
    samples. Error bars are standard error. Per the method described in\cite{tibshirani01gap}, 4 is the smallest $k$ such
    that $\text{Gap}(k) \geq \text{Gap}(k + 1) - \text{se}_{k + 1}$.}
  \label{fig:gap}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{stability-measures.pdf}
  \caption{Average proportion of non-overlap (APN), average distance (AD), average distance
    between means (ADM), and figure of merit (FOM) stability measures\cite{datta2006methods,
    yeung2001validating}. Statistics should be minimized; notice the local optima at $k = 4$ for
  APN, AD, and ADM.}
  \label{fig:stability}
\end{figure}

\subsubsection{Interpretation}

For the clustering on nonmotor domains, we displayed the distribution of each symptom for the four
clusters using boxplots, which allowed us to visualize the center and spread of each cluster. Since
the number of variables was larger for the clustering on the individual symptoms, we visualized
results for the second analysis with a heatmap. Finally, for the hierarchical clustering on the
symptoms themselves, we displayed results in a dendrogram and clustered the symptoms into four
interpretable clusters.

For each symptom in both clusterings, we used one-way ANOVA and $\chi^2$ tests to test the equality
of symptom means and proportions across the clusters found, using Bonferroni correction for
multiple testing with corrected $p < 0.05$ considered significant. Differences among pairwise
clusters were tested post-hoc using Tukey's range test for continuous means, or pairwise $\chi^2$
tests with Bonferroni correction for proportions.

\section{Results}

\subsection{Clustering with nonmotor domains}

$k$-means clustering on the nine nonmotor domains and the four motor symptoms found four clusters,
available in Table~\ref{tab:nmd} with boxplots in Figure~\ref{fig:box}. Cluster means for all
symptoms were found to be statistically different $(p < 0.05 / 13 \approx 0.003)$, with specific
pairwise differences noted in the table. Additionally, statistics for variables not used in the
clustering were explored similarly in Table~\ref{tab:nmd_extra}; cluster means for all variables
were found to be statistically different, except for PD onset.

Cluster 1 ($n = 406$) patients were mildly affected in all domains. This cluster was characterized
by relatively lower disease durations and ages.

Cluster 2 ($n = 189$) patients were severely affected in nonmotor domains but mildly affected in
motor domains. This cluster had an incidence of motor symptoms relatively similar to the Cluster 1
(mild) subtype especially in tremor, but generally expressed significantly higher scores for
nonmotor symptoms than Clusters 1 and 3, especially in the sleep/fatigue, mood/cognition and
miscellaneous domains.
% But also high variability

Cluster 3 ($n = 221$) patients were severely affected in motor domains but mildly affected in
nonmotor domains. Mean motor scores were greater than the means of clusters 1 and 2, but less than
4, with the exception of tremor, which was especially high.

Cluster 4 ($n = 88$) patients were severely affected in all domains, having the greatest symptom
mean out of all four clusters with the exception of tremor. Consequently, patients in Cluster 4 had
the longest average PD duration and oldest ages, but did not have a significantly different age of
PD onset.

\begin{table*}[t]
  \centering
  \caption{Summaries of clustering on nonmotor domains. Unless otherwise specified, statistics are
    reported as mean (sd). \textbf{Bold} indicates the highest score for that symptom out of the
  four clusters.}
  \label{tab:nmd}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    $n$ & 406 (45\%) & 189 (21\%) & 221 (24\%) & 88 (10\%) \\
    \midrule
    \textbf{Nonmotor} (Domains) & & & & \\
    Cardiovascular (0--24) & 0.7 (1.6)\tnote{24} & 2.3 (2.9)\tnote{134} & 1.1 (2.0)\tnote{24} &
    \textbf{7.0} (6.0)\tnote{234} \\
    Sleep/fatigue (0--48) & 4.3 (4.8)\tnote{234} & 14.9 (8.2)\tnote{134} & 6.9 (6.3)\tnote{124}
    & \textbf{21.0} (10.0)\tnote{123} \\
    Mood/cognition (0--60) & 3.1 (4.5)\tnote{234} & 16.1 (14.0)\tnote{134} & 6.6
    (8.1)\tnote{124} & \textbf{23.4} (14.3)\tnote{123} \\
    Perception/hallucination (0--36) & 0.5 (1.7)\tnote{24} & 1.9 (3.5)\tnote{134} & 0.8
    (2.0)\tnote{24} & \textbf{8.6} (6.9)\tnote{123} \\
    Attention/memory (0--36) & 2.8 (4.2)\tnote{24} & 8.9 (8.1)\tnote{134} & 3.2 (4.4)\tnote{24} &
    \textbf{15.6} (11.1)\tnote{123} \\
    Gastrointestinal (0--36) & 2.8 (3.9)\tnote{234} & 8.8 (7.1)\tnote{134} & 4.1
    (4.7)\tnote{124} & \textbf{14.6} (9.6)\tnote{123} \\
    Urinary (0--36) & 4.5 (5.9)\tnote{234} & 12.4 (9.7)\tnote{134} & 6.1 (6.4)\tnote{124} &
    \textbf{20.3} (9.9)\tnote{123} \\
    Sexual function (0--24) & 1.5 (3.1)\tnote{24} & 6.4 (7.4)\tnote{134} & 2.5 (4.1)\tnote{24} &
    \textbf{9.0} (9.7)\tnote{123} \\
    Miscellaneous (0--48) & 3.9 (4.7)\tnote{24} & 13.2 (8.6)\tnote{13} & 5.2 (5.8)\tnote{24} &
    \textbf{14.0} (9.5)\tnote{13} \\
    \midrule
    \textbf{Motor} & & & & \\
    Tremor & 1.9 (1.8)\tnote{34} & 1.6 (1.9)\tnote{34} & \textbf{4.3} (2.8)\tnote{124} & 3.5
    (3.8)\tnote{123} \\
    Bradykinesia & 1.5 (0.9)\tnote{234} & 2.1 (1.1)\tnote{134} & 3.6 (0.9)\tnote{124} &
    \textbf{4.0} (1.5)\tnote{123} \\
    Rigidity & 1.5 (0.9)\tnote{234} & 1.8 (1.1)\tnote{134} & 3.3 (0.9)\tnote{124} &
    \textbf{3.8} (1.4)\tnote{123} \\
    Axial & 1.8 (1.6)\tnote{234} & 3.3 (2.1)\tnote{134} & 4.3 (2.4)\tnote{124} &
    \textbf{7.4} (2.9)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
    \item[\textdagger] Another footnote.
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[b]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4-pub.pdf}
  \vspace{-7.5em}
  \captionsetup{justification=raggedleft,
    singlelinecheck=false
  }
  \caption{Boxplot summaries for clusters found \\ when clustering with nonmotor domains.}
  \label{fig:box}
\end{figure*}

\begin{table*}[t]
  \centering
  \caption{Cluster statistics for variables not used in the analysis}
  \label{tab:nmd_extra}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    \midrule
      Sex (\% male) & 63 & 53\tnote{3} & 71\tnote{2} & 58 \\
      Age & 62.7 (9.7)\tnote{34} & 64 (9.0)\tnote{4} & 65 (10.2)\tnote{14} & 70.5 (8.9)\tnote{123} \\
      PD\_Onset & 56 (10.6) & 55.2 (10.1) & 56.9 (11.1) & 58.2 (11.4) \\
      PD\_Duration & 6.7 (4.8)\tnote{234} & 8.7 (5.7)\tnote{14} & 8 (5.5)\tnote{14} & 12.3
      (8.1)\tnote{123} \\
      CISI\_Total & 5.7 (3.2)\tnote{234} & 9.2 (3.9)\tnote{14} & 9.7 (3.5)\tnote{14} & 14.8
      (5.1)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\subsection{Clustering with nonmotor symptoms}

$k$-means clustering performed on the 30 individual nonmotor symptoms found similar patterns to the
first clustering (Table~\ref{tab:nms30}, heatmap in Figure~\ref{fig:nms30-hm}). Means of all symptoms were found to
differ across clusters except for PD\_Onset and tremor.

Clusters 1 ($n = 509$) and 4 ($n = 49$), the mild and severe subtypes of the previous clustering,
retain their characteristics here, while clusters 2 ($n = 97$) and 3 ($n = 249$) are different. 2
is a mood/cognition-dominant cluster, with relatively higher proportion of female patients and
additionally a relatively more severe incidence of insomnia. 3 is not characterized by any
especially high symptoms and is thus an ``average'' group of patients.
% Once again, CISI TOTAL...

% Perhaps this table belongs in appendix?
\begin{table*}[t]
  \centering
  \caption{Summaries of clustering on individual nonmotor domains. Unless otherwise specified,
    statistics are reported as mean (sd). \textbf{Bold} indicates the highest score for that
  symptom out of the four clusters.}
  \label{tab:nms30}
  \begin{threeparttable}
  \small
\begin{tabular}{lrllll}
  \toprule
    & Cluster & 1 & 2 & 3 & 4 \\
    & $n$ & 509 (56.3\%) & 97 (10.7\%) & 249 (27.5\%) & 49 (5.4\%) \\
  \midrule
  1.\ Cardiovascular &
  Lightheadedness & 0.6 (1.3)\tnote{234} & 2.4 (3.0)\tnote{14} & 2.1 (2.9)\tnote{14} & \textbf{5.2}
  (4.2)\tnote{123} \\
  &Fainting & 0.1 (0.7)\tnote{24} & 0.5 (1.4)\tnote{134} & 0.2 (0.6)\tnote{24} & \textbf{2.4}
  (3.5)\tnote{123} \\
  \midrule
  2.\ Sleep/fatigue &
  Drowsiness & 1.0 (1.9)\tnote{234} & 2.8 (3.3)\tnote{14} & 2.5 (3.0)\tnote{14} & \textbf{7.0}
  (3.8)\tnote{123} \\
  &Fatigue & 1.3 (2.0)\tnote{234} & 6.3 (4.2)\tnote{13} & 4.8 (3.9)\tnote{124} & \textbf{7.3} (3.5)\tnote{13} \\
  &Insomnia & 1.2 (2.6)\tnote{234} & \textbf{5.5} (4.8)\tnote{13} & 3.2 (3.8)\tnote{12} & 4.2
  (3.9)\tnote{1} \\
  &RLS & 0.7 (1.6)\tnote{234} & 2.3 (3.5)\tnote{14} & 1.9 (3.2)\tnote{14} & \textbf{3.9}
  (3.9)\tnote{123} \\
  \midrule
  3.\ Mood/cognition &
  Loss\_interest & 0.4 (1.1)\tnote{234} & \textbf{6.1} (3.8)\tnote{134} & 1.0 (1.8)\tnote{124} & 4.6 (3.7)\tnote{123} \\
  &Loss\_activities & 0.7 (1.6)\tnote{234} & \textbf{7.1} (4.1)\tnote{134} & 1.7 (2.4)\tnote{124} &
  5.4 (4.0)\tnote{123} \\
  &Anxiety & 0.9 (1.8)\tnote{234} & \textbf{6.6} (4.3)\tnote{134} & 2.0 (2.8)\tnote{124} & 4.6 (3.9)\tnote{123} \\
  &Depression & 0.8 (1.6)\tnote{234} & \textbf{7.8} (3.7)\tnote{134} & 2.3 (3.0)\tnote{124} & 4.3 (3.7)\tnote{123} \\
  &Flat\_affect & 0.4 (1.1)\tnote{234} & \textbf{4.9} (4.3)\tnote{134} & 0.9 (1.9)\tnote{124} & 2.7 (3.0)\tnote{123} \\
  &Loss\_pleasure & 0.4 (1.2)\tnote{234} & \textbf{6.8} (4.1)\tnote{134} & 1.1 (2.1)\tnote{124} & 3.4 (3.3)\tnote{123} \\
  \midrule
  % This one is spread across multiple cells
  4.\ Perception/ & Hallucination & 0.2 (1.0)\tnote{234} & 0.8 (1.9)\tnote{14} & 0.6 (1.6)\tnote{14} & \textbf{4.6} (3.6)\tnote{123} \\
  hallucination &Delusion & 0.1 (0.7)\tnote{24} & 1.5 (3.1)\tnote{124} & 0.2 (1.0)\tnote{24} &
  \textbf{3.1} (3.8)\tnote{123} \\
  &Diplopia & 0.2 (0.7)\tnote{234} & 1.0 (2.5)\tnote{14} & 0.6 (1.4)\tnote{14} & \textbf{4.4}
  (4.3)\tnote{123} \\
  \midrule
  5.\ Attention/ & Loss\_concentration & 0.9 (1.8)\tnote{234} & 4.0 (3.6)\tnote{134} & 2.5 (3.1)\tnote{124} & \textbf{6.6} (3.9)\tnote{123} \\
  memory &Forget\_explicit & 0.9 (1.6)\tnote{234} & 3.8 (3.9)\tnote{134} & 2.2 (2.9)\tnote{124} & \textbf{6.7} (3.6)\tnote{123} \\
  &Forget\_implicit & 0.7 (1.6)\tnote{234} & 3.1 (3.7)\tnote{134} & 1.8 (2.6)\tnote{124} & \textbf{6.7} (4.0)\tnote{123} \\
  \midrule
  6.\ Gastrointestinal &
  Drooling & 0.7 (1.6)\tnote{234} & 3.0 (4.1)\tnote{14} & 3.0 (3.8)\tnote{14} & \textbf{6.1} (4.4)\tnote{123} \\
  &Swallowing & 0.4 (1.2)\tnote{234} & 1.2 (2.1)\tnote{14} & 1.7 (2.8)\tnote{14} & \textbf{4.5} (3.9)\tnote{123} \\
  &Constipation & 1.6 (2.9)\tnote{234} & 3.6 (4.3)\tnote{14} & 3.3 (4.1)\tnote{14} & \textbf{7.0} (5.1)\tnote{123} \\
  \midrule
  7.\ Urinary &
  Urinary\_urgency & 1.0 (1.8)\tnote{234} & 3.5 (4.1)\tnote{14} & 3.8 (3.8)\tnote{14} & \textbf{7.8} (4.0)\tnote{123} \\
  &Urinary\_frequency & 0.9 (1.9)\tnote{234} & 4.0 (4.1)\tnote{14} & 3.8 (4.0)\tnote{14} & \textbf{6.5} (4.0)\tnote{123} \\
  &Nocturia & 1.7 (2.3)\tnote{234} & 5.0 (4.3)\tnote{14} & 5.1 (4.1)\tnote{14} & \textbf{7.7} (3.9)\tnote{123} \\
  \midrule
  8.\ Sexual &
  Sex\_drive & 0.8 (1.9)\tnote{234} & 4.5 (4.9)\tnote{13} & 2.3 (3.6)\tnote{124} & \textbf{5.2} (5.4)\tnote{13} \\
  &Sex\_dysfunction & 0.8 (2.2)\tnote{234} & 3.2 (4.7)\tnote{1} & 2.4 (3.8)\tnote{14} &
  \textbf{4.6} (5.4)\tnote{13} \\
  \midrule
  9.\ Miscellaneous &
  Unexplained\_pain & 0.7 (1.9)\tnote{234} & 2.5 (3.7)\tnote{14} & 3.0 (4.1)\tnote{14} & \textbf{4.2} (4.3)\tnote{123} \\
  &Gustation\_olfaction & 1.5 (2.9)\tnote{234} & 3.6 (4.3)\tnote{1} & 3.2 (3.9)\tnote{14} &
  \textbf{4.7} (4.8)\tnote{13} \\
  &Weight\_change & 0.9 (1.7)\tnote{234} & 2.5 (3.4)\tnote{14} & 1.9 (3.0)\tnote{14} & \textbf{3.9} (4.2)\tnote{123} \\
  &Sweating & 0.7 (1.8)\tnote{234} & 2.2 (3.5)\tnote{1} & 2.8 (4.1)\tnote{1} & \textbf{3.2}
  (3.8)\tnote{1} \\
  \midrule
  Motor &
  Tremor & 2.5 (2.3) & 2.1 (2.4) & 2.6 (2.6) & \textbf{4.1} (4.3) \\
  &Bradykinesia & 2.0 (1.2)\tnote{234} & 2.8 (1.4)\tnote{14} & 2.8 (1.4)\tnote{14} & \textbf{3.9} (1.8)\tnote{123} \\
  &Rigidity & 1.9 (1.2)\tnote{234} & 2.4 (1.3)\tnote{14} & 2.6 (1.4)\tnote{14} & \textbf{3.6} (1.6)\tnote{123} \\
  &Axial & 2.2 (1.8)\tnote{234} & 4.4 (2.8)\tnote{14} & 4.2 (2.7)\tnote{14} & \textbf{7.2} (3.2)\tnote{123} \\
  \midrule
  Variables not & Age & 62.7 (9.9)\tnote{34} & 64.5 (9.6)\tnote{4} & 66 (9.2)\tnote{14} & 71.5
  (8.1)\tnote{123} \\
  used in analysis & Sex (\% Male) & 65\tnote{2} & 51\tnote{1} & 60 & 63 \\
  & PD\_Onset & 55.9 (10.8) & 56.2 (10.4) & 56.7 (10.5) & 57.8 (11.0) \\
  & PD\_Duration & 6.8 (4.8)\tnote{34} & 8.3 (5.3)\tnote{4} & 9.3 (6.3)\tnote{14} & 13.8
  (8.2)\tnote{123} \\
  & CISI\_Total & 6.3 (3.4)\tnote{234} & 10.8 (4.7)\tnote{14} & 10 (4.0)\tnote{14} & 15.1 (5.7)\tnote{123} \\
   \bottomrule
\end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{nms30m-colhc-pub.pdf}
  \caption{Hierarchical clustering of motor (blue) and nonmotor (black) symptoms. Symptoms are
  labeled with their name and corresponding domain number. Dendrogram colored with 4 clusters.}
  \label{fig:hc}
\end{figure*}

% TODO: If the outline of the paper is good enough it might be possible to have this inline
% in the paper somewhere. If so, change figure* -> figure, make h! (force horizontal?), and switch
% 0.8 linewidth to full linewidth
% Also, I'm going to include this AFTER nms30m-colhc, so that colhc/boxplots fits nicely on one
% page
\begin{figure*}[p]
  \centering
  \includegraphics[width=0.7\linewidth]{nms30-hm-pub.pdf}
  % TODO: Variables not used in analysis? Label color - black nms, blue nonmotor, grey ?
  \caption{Heatmap of clusters found when clustering with nonmotor symptoms. Color indicates
    cluster mean as a $z$-score for the given symptom; red is higher, green is lower.}
  \label{fig:nms30-hm}
\end{figure*}

\subsection{Hierarchical clustering on symptoms}

Hierarchical clustering on the 30 nonmotor symptoms and the four motor symptoms produced the
dendrogram in Figure~\ref{fig:hc}. Predictably, symptoms in the same nonmotor domains tended to
cluster together, with the exceptions of diplopia (domain 4), drowsiness (domain 2), and the
symptoms in domain 9 (miscellaneous). Notably, tremor was the most dissimilar symptom, occupying a
single branch at the top of the tree.

\section{Discussion}

$k$-means clustering on this dataset reveals clusters that confirm previous
findings in the field, mainly van Rooden et al.\cite{vanrooden10} and the identification of four
subtypes of Parkinson's disease: mild, nonmotor-predominant, motor-predominant, and severe.  van
Rooden's work was done with a separate dataset using a different modeling method
(expectation-maximization), and this investigation independently confirms these subtype
classifications. Unlike van Rooden, mean disease durations differences do exist between subtypes 1
(mild) and 4 (severe), likely due to further development of the disease, although the differences
between 2 and 3 (nonmotor/motor predominated) subtypes are insignificant
(Table~\ref{tab:tukeyhsd}), suggesting different developmental paths of the disease.

Overall, little information was found in pdonset,
durat\_pd, or current age, according to Tables~\ref{tab:tukeyhsd}
and~\ref{tab:info_gain}.  Mean ages were similar for subgroups 1, 2, and 3 ($p >
0.05$), but different for the severe subtype 4, which makes sense given that
patients in 4 also have longer disease durations. Specifically, clusters 1 and
4 seem to be phenotypically quite similar, except at different stages of
disease progression, given cluster 4's higher age and durat\_pd scores.

However, clusters 2 and 3 clearly show different disease progression, one in
the motor direction, and one in the nonmotor. Both groups have similar age,
pdonset, and durat\_pd scores, but differ wildly in symptomatic expression.
Cluster 2 is dominated by a high prevalence of nonmotor symptoms, such as
nms\_d2, nms\_d3, nms\_d7, and nms\_d9. Cluster 3, however, is dominated by a
high prevalence of motor symptoms, while most motor symptoms are similar to the
mild cluster 1. Of note is that the tremor population mean
is the highest cluster mean, even higher than the severe subtype 4. This
motor-dominant cluster may thus overlap with Ma's tremor dominant/slow
progression cluster\cite{ma15}.

Generally, given stable pdonset scores and predictably increasing durat\_pd
scores for clusters 1 and 4, Ma et al's rapid disease
progression/late onset and tremor dominant/slow progression clusters
\cite{ma15} were mostly not found in this dataset, save for the tremor-dominant
motor cluster.

The most important nonmotor symptoms in determining these clusters were nms\_d2
(sleep) and nms\_d3 (mood/cognition), which echo findings of Fereshtehnejad's
longitudinal study\cite{fereshtehnejad15} and are similar to Sauerbier's
identification of sleep dominant and cognitive dominant clinical NMS subtypes\cite{sauerbier15}.
Compared to Erro et al.\cite{erro13}, nonmotor/motor
dominant subtypes were indeed found, but an additional subgroup with relatively
severe levels of both motor and nonmotor symptoms were found. Erro's benign
subtype groups possibly overlap with the mild cluster 1 found in this
investigation.

\subsection{Nonmotor subtype: clustering and modeling}
Nonmotor symptoms nms\_d2 and nms\_d3 became critical not only in
classification trees distinguishing between the various symptoms but in the
nonmotor-predominant subgroup itself. In $k$-means subdivision of the
nonmotor-dominant subtype where $k = 2$ and $k = 3$, opposite trends were
confirmed with nms\_d2 and nms\_d3 symptoms. Similarly, in the 2 and 4 vs rest
decision tree (Figure~\ref{fig:dtree-2and4va-pruned}), nms\_d2 and nms\_d3 nodes were
used to differentiate various categories of nomotor-dominant patients.
When $k = 3$, the subtype with the highest nms\_d2 scores and lowest
nms\_d3 scores had by far the highest axial scores, nms\_d6 (gastrointestinal)
scores, and nms\_d7 (urinary) scores. Thus subtype 3 of the nonmotor-dominated
group could include patients falling into the cognitive/depression-dominant or
autonomic dominant subtypes.

Despite the variety in symptomatic expression in this nonmotor group, what
seems most consistent is the presence of nms\_d9 (miscellaneous) nonmotor
symptoms, as it is used as the root node of the 2 vs all decision tree
(Figure~\ref{fig:2va}) and the 2 and 4 vs rest decision tree
(Figure~\ref{fig:dtree-2and4va-pruned}).

First, the longitudinal analysis gives more insight into the clusters found when clustering on
nms\_d\{1-9\}. According to Figure~\ref{fig:longcorr}, most symptoms are highly correlated, with PD
duration, but notably, mood/cognition symptoms (nms9, nms10, nms12) and tremor are not correlated
highly with PD duration. The differences in disease progression can be seen by the corresponding
graphs, Figures~\ref{fig:nms9-multi} and~\ref{fig:nms10-multi}. In both graphs, what is interesting
is that Subtype 2 (Nonmotor-Dominant) starts at higher scores for nms\_9 and nms\_10, thus
indicating that these patients' subtype is can be determined early in PD duration from depressive
symptom score. Similarly, when examining Subtype 3 (Motor-dominant) in
Figure~\ref{fig:tremor-multi}, the mean tremor score is substantially higher from PD onset.
Interestingly, Subtype 4 (Severe) generally starts at lower tremor and motor scores during disease
onset (Figure~\ref{fig:cisitot-multi}), but then rises sharply, exceeding other Subtypes. More evidence
that tremor is a unique motor symptom is located in Figure~\ref{fig:nms30m-colhc}, where it is the most
distant symptom from all other symptoms.

When examining all 30 symptoms, more evidence is given that the previously-discovered Subtype
2 (Nonmotor-Dominant) may be primarily characterized by high depressive symptoms. By graphing the
30 subtypes attached to the original clustering, as in Figure~\ref{fig:d9to30}, the mean scores of
nms8, nms9, and nms10 for Subtype 2 are substantially higher than Subtypes 1 and 3.

Indeed, PCA on the 30 nonmotor symptoms identifies the second-most prominent component as a general
mood/cognition component, and $k$-means clustering on the 30 symptoms only
(Figure~\ref{fig:nms30m-k4-heatmap-improved}) divides the 1000 patients into four slightly
different groups, a mild, average, depression-dominant, and severe group.

The Gaussian mixture model identified in Figure~\ref{fig:nms30m-mclust-heatmap-improved} fragments
the previously-discovered clusters into more groups. Here, a wide variety of specialized subtypes
of PD are displayed, including insomnia, urinary, motor, nonmotor, and depression-dominant groups,
as well as the expected mild, average, and severe subtypes. It is likely that the previous analysis
with only nms\_d\{1-9\} combined most of those specialized groups into the more general
nonmotor-dominant Subtype 2.

It's intuitive that a Depression-Dominant group emerges when clustering on nms\{1-30\}, since
domain 3 consists of 5 separate measures. Thus, any high expression of depressive symptoms is
magnified in clustering, since the symptoms are highly similar (Figure~\ref{fig:nms30m-colhc}) and
treated with equal weight. Once again reinforcing what was discovered previously, depressive
symptoms have been shown to be very important in determining subtypes of PD.

\section{Supplementary Material}
\section{References}

\bibliography{manuscript}

\end{document}