### 2016-05-09 writeup/manuscript/manuscript.tex
Updated analyses and draft of materials/results

% PREAMBLE ====
\documentclass[preprint,5p]{elsarticle} % TODO: What model?

% Use times font
\usepackage{mathptmx}
\usepackage{lmodern}

% Math
\usepackage{amsmath}

% Tables, Figures
\usepackage{booktabs}
\usepackage{threeparttable}
% For line-breaking cells
\newcommand{\spc}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\usepackage{float}
\usepackage{caption}
\setlength{\tabcolsep}{10pt}

% Graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png}
\graphicspath{{../../figures/}}

% References
\usepackage{numcompress}
\biboptions{super}
\bibliographystyle{model3-num-names}

% DOCUMENT ====
\begin{document}
% FRONTMATTER ====
\begin{frontmatter}
\title{Parkinson's disease cluster analysis}
% \author[c3]{Pablo Martinez-Martin}
% \ead{pmartinez@isciii.es}
% \author[bc]{Jesse Mu}
% \ead{muj@bc.edu}
% \author[cig]{Concha Bielza}
% \ead{mcbielza@fi.upm.es}
% \author[cig]{Pedro Larra\~naga}
% \ead{plarranaga@fi.upm.es}
% \address[c3]{Area of Applied Epidemiology, National Centre of Epidemiology and CIBERNED, Carlos III Institute of Health, Madrid, Spain}
% \address[bc]{Department of Computer Science, Boston College, Chestnut Hill, Massachusetts, USA}
% \address[cig]{Computational Intelligence Group, Polytechnic University of Madrid, Madrid, Spain}

\begin{abstract}
\emph{Background} \quad This is the background. \\
\emph{Methods} \quad  These are methods. \\
\emph{Findings} \quad These are findings. \\
\emph{Interpretation} \quad This is interpretation. \\
\emph{Funding} \quad This is funding.
\end{abstract}
\journal{The Lancet Neurology}
\end{frontmatter}

% BODY ====

\section{Introduction}

\section{Methods}

\subsection{Statistical analysis}

Out of the 951 patients in the study, we used listwise deletion to exclude 47 patients due to
missing measurements in the variables of interest, resulting in 904 remaining patients. All
symptoms were scaled to $z$-scores ($\mu = 0$, $\sigma = 1$) before clustering, and unscaled
afterwards for interpretation. Analyses were conducted in R 3.2.4 (www.r-project.org).

\subsubsection{Cluster analysis}

$k$-means was used for cluster analysis. We performed two analyses on the
patients in the dataset: the first on the nine aggregate nonmotor symptom
domains and the four motor symptoms (``domains clustering''), and the second on the 30 individual
nonmotor symptoms collected with the four motor symptoms (``symptoms clustering''). Complete-linkage
hierarchical clustering on the 30 nonmotor symptoms was also performed to observe the
categorization of symptoms.

\subsubsection{Determining $k$}

Various formal measures were used to determine the optimal number of
clusters for the dataset. First, the optimal $k$ according to the Gap Statistic and the
1-standard-error method\cite{tibshirani01gap} was $k = 4$ (Figure~\ref{fig:gap}). Second, although
the globally-optimal clusters reported from stability measures of the clValid R package
\cite{brock11clvalid} stability were 2, 6, and 8 (Figure~\ref{fig:stability}), there appeared to be
consensus that $k = 4$ is a locally-optimal number of clusters which supported the gap statistic
result.

Other cluster validation methods (SSE scree plot, minimum average silhouette width) generally
suggested 2 clusters. As a consequence, $k = 2, 3, 4$ was tried, where $k = 2, 3$ simply divided
the data uninformatively into groups with varying levels of overall PD severity. Thus $k = 4$ was
selected to offer a good blend of model fit, informativeness, and parsimony.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.pdf}
  \caption{Plot of gap statistic versus number of clusters with $k$-means on 100 bootstrapped
    samples. Error bars are standard error. Per the method described in Tibshirani\cite{tibshirani01gap}, 4 is the smallest $k$ such
    that $\text{Gap}(k) \geq \text{Gap}(k + 1) - \text{se}_{k + 1}$.}
  \label{fig:gap}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{stability-measures.pdf}
  \caption{Average proportion of non-overlap (APN), average distance (AD), average distance
    between means (ADM), and figure of merit (FOM) stability measures\cite{datta2006methods,
    yeung2001validating}. Statistics should be minimized; notice the local optima at $k = 4$ for
  APN, AD, and ADM.}
  \label{fig:stability}
\end{figure}

\subsubsection{Interpretation}

For the domains clustering, we displayed the distribution of each symptom for the four
clusters using boxplots, which allowed us to visualize the center and spread of each cluster. Since
the number of variables was larger for the symptoms clustering, we visualized results for the
second analysis with a heatmap. Finally, for the hierarchical clustering on the symptoms
themselves, we displayed results in a dendrogram and clustered the symptoms into four interpretable
clusters.

For each symptom in both clusterings, we used one-way ANOVA and $\chi^2$ tests to test the equality
of symptom means and proportions across the clusters found, using Bonferroni correction for
multiple testing with corrected $p < 0.05$ considered significant. Differences among pairwise
clusters were tested post-hoc using Tukey's range test for continuous means, or pairwise $\chi^2$
tests with Bonferroni correction for proportions.

To compare the clusterings, we computed the adjusted rand index\cite{hubert1985} (ARI) and
$v$-measure\cite{rosenberg07vmeasure} to evaluate similarity between the two clusterings, and used
stacked barplots to visualize the alignment of the clusters.

Lastly, to explore the relationship between symptom severity and disease duration, we computed the
linear correlation for each symptom with PD duration and fitted smoothed loess curves to the
data both globally and for each cluster found in the previous analyses.

\section{Results}

\subsection{Domains clustering}

$k$-means clustering on the nine nonmotor domains and the four motor symptoms found four clusters
(Table~\ref{tab:nmd}; boxplots in Figure~\ref{fig:box}). Cluster means for all
symptoms were found to be statistically different $(p < 0.05 / 13 \approx 0.003)$, with specific
pairwise differences noted in the table. Additionally, statistics for variables not used in the
clustering were explored similarly in Table~\ref{tab:nmd_extra}; cluster means for all variables
were found to be statistically different, except for PD onset.

Cluster 1 ($n = 406$) patients were mildly affected in all domains. This cluster was characterized
by relatively lower disease durations and ages.

Cluster 2 ($n = 189$) patients were severely affected in nonmotor domains but mildly affected in
motor domains. This cluster had an incidence of motor symptoms relatively similar to the Cluster 1
(mild) subtype especially in tremor, but generally expressed significantly higher scores for
nonmotor symptoms than Clusters 1 and 3, especially in the sleep/fatigue, mood/cognition and
miscellaneous domains. This group also had a statistically higher percentage of females than
Cluster 3.
% But also high variability

Cluster 3 ($n = 221$) patients were severely affected in motor domains but mildly affected in
nonmotor domains. Mean motor scores were greater than the means of clusters 1 and 2, but less than
4, with the exception of tremor, which was especially high.

Cluster 4 ($n = 88$) patients were severely affected in all domains, having the greatest symptom
mean out of all four clusters with the exception of tremor. Consequently, patients in Cluster 4 had
the longest average PD duration and oldest ages, but did not have a significantly different age of
PD onset.

\begin{table*}[t]
  \centering
  \caption{Summaries of domains clustering. Unless otherwise specified, statistics are
    reported as mean (sd). \textbf{Bold} indicates the highest score for that symptom out of the
  four clusters.}
  \label{tab:nmd}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    $n$ & 406 (45\%) & 189 (21\%) & 221 (24\%) & 88 (10\%) \\
    \midrule
    \textbf{Nonmotor} (Domains) & & & & \\
    Cardiovascular (0--24) & 0.7 (1.6)\tnote{24} & 2.3 (2.9)\tnote{134} & 1.1 (2.0)\tnote{24} &
    \textbf{7.0} (6.0)\tnote{234} \\
    Sleep/fatigue (0--48) & 4.3 (4.8)\tnote{234} & 14.9 (8.2)\tnote{134} & 6.9 (6.3)\tnote{124}
    & \textbf{21.0} (10.0)\tnote{123} \\
    Mood/cognition (0--60) & 3.1 (4.5)\tnote{234} & 16.1 (14.0)\tnote{134} & 6.6
    (8.1)\tnote{124} & \textbf{23.4} (14.3)\tnote{123} \\
    Perception/hallucination (0--36) & 0.5 (1.7)\tnote{24} & 1.9 (3.5)\tnote{134} & 0.8
    (2.0)\tnote{24} & \textbf{8.6} (6.9)\tnote{123} \\
    Attention/memory (0--36) & 2.8 (4.2)\tnote{24} & 8.9 (8.1)\tnote{134} & 3.2 (4.4)\tnote{24} &
    \textbf{15.6} (11.1)\tnote{123} \\
    Gastrointestinal (0--36) & 2.8 (3.9)\tnote{234} & 8.8 (7.1)\tnote{134} & 4.1
    (4.7)\tnote{124} & \textbf{14.6} (9.6)\tnote{123} \\
    Urinary (0--36) & 4.5 (5.9)\tnote{234} & 12.4 (9.7)\tnote{134} & 6.1 (6.4)\tnote{124} &
    \textbf{20.3} (9.9)\tnote{123} \\
    Sexual function (0--24) & 1.5 (3.1)\tnote{24} & 6.4 (7.4)\tnote{134} & 2.5 (4.1)\tnote{24} &
    \textbf{9.0} (9.7)\tnote{123} \\
    Miscellaneous (0--48) & 3.9 (4.7)\tnote{24} & 13.2 (8.6)\tnote{13} & 5.2 (5.8)\tnote{24} &
    \textbf{14.0} (9.5)\tnote{13} \\
    \midrule
    \textbf{Motor} & & & & \\
    Tremor & 1.9 (1.8)\tnote{34} & 1.6 (1.9)\tnote{34} & \textbf{4.3} (2.8)\tnote{124} & 3.5
    (3.8)\tnote{123} \\
    Bradykinesia & 1.5 (0.9)\tnote{234} & 2.1 (1.1)\tnote{134} & 3.6 (0.9)\tnote{124} &
    \textbf{4.0} (1.5)\tnote{123} \\
    Rigidity & 1.5 (0.9)\tnote{234} & 1.8 (1.1)\tnote{134} & 3.3 (0.9)\tnote{124} &
    \textbf{3.8} (1.4)\tnote{123} \\
    Axial & 1.8 (1.6)\tnote{234} & 3.3 (2.1)\tnote{134} & 4.3 (2.4)\tnote{124} &
    \textbf{7.4} (2.9)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[b]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4-pub.pdf}
  \vspace{-7.5em}
  \captionsetup{justification=raggedleft,
    singlelinecheck=false
  }
  \caption{Boxplot cluster summaries \\ for domains clustering.}
  \label{fig:box}
\end{figure*}

\begin{table*}[t]
  \centering
  \caption{Cluster statistics for variables not used in the analysis}
  \label{tab:nmd_extra}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    \midrule
      Sex (\% male) & 63 & 53\tnote{3} & 71\tnote{2} & 58 \\
      Age & 62.7 (9.7)\tnote{34} & 64 (9.0)\tnote{4} & 65 (10.2)\tnote{14} & 70.5 (8.9)\tnote{123} \\
      PD\_Onset & 56 (10.6) & 55.2 (10.1) & 56.9 (11.1) & 58.2 (11.4) \\
      PD\_Duration & 6.7 (4.8)\tnote{234} & 8.7 (5.7)\tnote{14} & 8 (5.5)\tnote{14} & 12.3
      (8.1)\tnote{123} \\
      CISI\_Total & 5.7 (3.2)\tnote{234} & 9.2 (3.9)\tnote{14} & 9.7 (3.5)\tnote{14} & 14.8
      (5.1)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\subsection{Symptoms clustering}

$k$-means clustering performed on the 30 individual nonmotor symptoms found similar patterns to the
first clustering (Table~\ref{tab:nms30}, heatmap in Figure~\ref{fig:nms30-hm}). Means of all
symptoms were found to differ across clusters except for PD\_Onset and tremor.

Clusters 1 ($n = 509$) and 4 ($n = 49$), the mild and severe subtypes of the previous clustering,
retained their characteristics here, while clusters 2 ($n = 97$) and 3 ($n = 249$) were different.
2 was a mood/cognition-dominant cluster, with a relatively higher proportion of female patients and
additionally a more severe incidence of insomnia. 3 was not characterized by any especially high
symptoms and was thus classified as a group of patients with average severity.
% Once again, CISI TOTAL...

\subsection{Comparison between clusterings}

The two clusterings are somewhat similar ($\text{ARI} = 0.46$; $\text{$v$-measure} = 0.43$).
Alignment of the clusters is visualized in Figure~\ref{fig:align}. While clusters 1 and 4 of both
clusterings align well with each other, main differences in the clusters occur in the mixing of
clusters 2 and 3 of both clusterings; for example, domains cluster 3 (motor-dominant) is somewhat
evenly split among symptoms clusters 1 (mild) and 3 (average).

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{cluster-alignment.pdf}
  \caption{Bar chart indicating the distribution of opposite cluster membership for the patients of
  a given cluster. For example, the bar furthest left indicates that close to 100\% of patients in
domains cluster 1 are in symptoms cluster 1, with the rest in symptoms cluster 3.}
  \label{fig:align}
\end{figure}

% Perhaps this table belongs in appendix?
\begin{table*}[t]
  \centering
  \caption{Summaries of clustering on individual nonmotor domains. Unless otherwise specified,
    statistics are reported as mean (sd). \textbf{Bold} indicates the highest score for that
  symptom out of the four clusters.}
  \label{tab:nms30}
  \begin{threeparttable}
  \small
\begin{tabular}{lrllll}
  \toprule
    & Cluster & 1 & 2 & 3 & 4 \\
    & $n$ & 509 (56.3\%) & 97 (10.7\%) & 249 (27.5\%) & 49 (5.4\%) \\
  \midrule
  1.\ Cardiovascular &
  Lightheadedness & 0.6 (1.3)\tnote{234} & 2.4 (3.0)\tnote{14} & 2.1 (2.9)\tnote{14} & \textbf{5.2}
  (4.2)\tnote{123} \\
  &Fainting & 0.1 (0.7)\tnote{24} & 0.5 (1.4)\tnote{134} & 0.2 (0.6)\tnote{24} & \textbf{2.4}
  (3.5)\tnote{123} \\
  \midrule
  2.\ Sleep/fatigue &
  Drowsiness & 1.0 (1.9)\tnote{234} & 2.8 (3.3)\tnote{14} & 2.5 (3.0)\tnote{14} & \textbf{7.0}
  (3.8)\tnote{123} \\
  &Fatigue & 1.3 (2.0)\tnote{234} & 6.3 (4.2)\tnote{13} & 4.8 (3.9)\tnote{124} & \textbf{7.3} (3.5)\tnote{13} \\
  &Insomnia & 1.2 (2.6)\tnote{234} & \textbf{5.5} (4.8)\tnote{13} & 3.2 (3.8)\tnote{12} & 4.2
  (3.9)\tnote{1} \\
  &RLS & 0.7 (1.6)\tnote{234} & 2.3 (3.5)\tnote{14} & 1.9 (3.2)\tnote{14} & \textbf{3.9}
  (3.9)\tnote{123} \\
  \midrule
  3.\ Mood/cognition &
  Loss\_interest & 0.4 (1.1)\tnote{234} & \textbf{6.1} (3.8)\tnote{134} & 1.0 (1.8)\tnote{124} & 4.6 (3.7)\tnote{123} \\
  &Loss\_activities & 0.7 (1.6)\tnote{234} & \textbf{7.1} (4.1)\tnote{134} & 1.7 (2.4)\tnote{124} &
  5.4 (4.0)\tnote{123} \\
  &Anxiety & 0.9 (1.8)\tnote{234} & \textbf{6.6} (4.3)\tnote{134} & 2.0 (2.8)\tnote{124} & 4.6 (3.9)\tnote{123} \\
  &Depression & 0.8 (1.6)\tnote{234} & \textbf{7.8} (3.7)\tnote{134} & 2.3 (3.0)\tnote{124} & 4.3 (3.7)\tnote{123} \\
  &Flat\_affect & 0.4 (1.1)\tnote{234} & \textbf{4.9} (4.3)\tnote{134} & 0.9 (1.9)\tnote{124} & 2.7 (3.0)\tnote{123} \\
  &Loss\_pleasure & 0.4 (1.2)\tnote{234} & \textbf{6.8} (4.1)\tnote{134} & 1.1 (2.1)\tnote{124} & 3.4 (3.3)\tnote{123} \\
  \midrule
  % This one is spread across multiple cells
  4.\ Perception/ & Hallucination & 0.2 (1.0)\tnote{234} & 0.8 (1.9)\tnote{14} & 0.6 (1.6)\tnote{14} & \textbf{4.6} (3.6)\tnote{123} \\
  hallucination &Delusion & 0.1 (0.7)\tnote{24} & 1.5 (3.1)\tnote{124} & 0.2 (1.0)\tnote{24} &
  \textbf{3.1} (3.8)\tnote{123} \\
  &Diplopia & 0.2 (0.7)\tnote{234} & 1.0 (2.5)\tnote{14} & 0.6 (1.4)\tnote{14} & \textbf{4.4}
  (4.3)\tnote{123} \\
  \midrule
  5.\ Attention/ & Loss\_concentration & 0.9 (1.8)\tnote{234} & 4.0 (3.6)\tnote{134} & 2.5 (3.1)\tnote{124} & \textbf{6.6} (3.9)\tnote{123} \\
  memory &Forget\_explicit & 0.9 (1.6)\tnote{234} & 3.8 (3.9)\tnote{134} & 2.2 (2.9)\tnote{124} & \textbf{6.7} (3.6)\tnote{123} \\
  &Forget\_implicit & 0.7 (1.6)\tnote{234} & 3.1 (3.7)\tnote{134} & 1.8 (2.6)\tnote{124} & \textbf{6.7} (4.0)\tnote{123} \\
  \midrule
  6.\ Gastrointestinal &
  Drooling & 0.7 (1.6)\tnote{234} & 3.0 (4.1)\tnote{14} & 3.0 (3.8)\tnote{14} & \textbf{6.1} (4.4)\tnote{123} \\
  &Swallowing & 0.4 (1.2)\tnote{234} & 1.2 (2.1)\tnote{14} & 1.7 (2.8)\tnote{14} & \textbf{4.5} (3.9)\tnote{123} \\
  &Constipation & 1.6 (2.9)\tnote{234} & 3.6 (4.3)\tnote{14} & 3.3 (4.1)\tnote{14} & \textbf{7.0} (5.1)\tnote{123} \\
  \midrule
  7.\ Urinary &
  Urinary\_urgency & 1.0 (1.8)\tnote{234} & 3.5 (4.1)\tnote{14} & 3.8 (3.8)\tnote{14} & \textbf{7.8} (4.0)\tnote{123} \\
  &Urinary\_frequency & 0.9 (1.9)\tnote{234} & 4.0 (4.1)\tnote{14} & 3.8 (4.0)\tnote{14} & \textbf{6.5} (4.0)\tnote{123} \\
  &Nocturia & 1.7 (2.3)\tnote{234} & 5.0 (4.3)\tnote{14} & 5.1 (4.1)\tnote{14} & \textbf{7.7} (3.9)\tnote{123} \\
  \midrule
  8.\ Sexual &
  Sex\_drive & 0.8 (1.9)\tnote{234} & 4.5 (4.9)\tnote{13} & 2.3 (3.6)\tnote{124} & \textbf{5.2} (5.4)\tnote{13} \\
  &Sex\_dysfunction & 0.8 (2.2)\tnote{234} & 3.2 (4.7)\tnote{1} & 2.4 (3.8)\tnote{14} &
  \textbf{4.6} (5.4)\tnote{13} \\
  \midrule
  9.\ Miscellaneous &
  Unexplained\_pain & 0.7 (1.9)\tnote{234} & 2.5 (3.7)\tnote{14} & 3.0 (4.1)\tnote{14} & \textbf{4.2} (4.3)\tnote{123} \\
  &Gustation\_olfaction & 1.5 (2.9)\tnote{234} & 3.6 (4.3)\tnote{1} & 3.2 (3.9)\tnote{14} &
  \textbf{4.7} (4.8)\tnote{13} \\
  &Weight\_change & 0.9 (1.7)\tnote{234} & 2.5 (3.4)\tnote{14} & 1.9 (3.0)\tnote{14} & \textbf{3.9} (4.2)\tnote{123} \\
  &Sweating & 0.7 (1.8)\tnote{234} & 2.2 (3.5)\tnote{1} & 2.8 (4.1)\tnote{1} & \textbf{3.2}
  (3.8)\tnote{1} \\
  \midrule
  Motor &
  Tremor & 2.5 (2.3) & 2.1 (2.4) & 2.6 (2.6) & \textbf{4.1} (4.3) \\
  &Bradykinesia & 2.0 (1.2)\tnote{234} & 2.8 (1.4)\tnote{14} & 2.8 (1.4)\tnote{14} & \textbf{3.9} (1.8)\tnote{123} \\
  &Rigidity & 1.9 (1.2)\tnote{234} & 2.4 (1.3)\tnote{14} & 2.6 (1.4)\tnote{14} & \textbf{3.6} (1.6)\tnote{123} \\
  &Axial & 2.2 (1.8)\tnote{234} & 4.4 (2.8)\tnote{14} & 4.2 (2.7)\tnote{14} & \textbf{7.2} (3.2)\tnote{123} \\
  \midrule
  Variables not & Age & 62.7 (9.9)\tnote{34} & 64.5 (9.6)\tnote{4} & 66 (9.2)\tnote{14} & 71.5
  (8.1)\tnote{123} \\
  used in analysis & Sex (\% Male) & 65\tnote{2} & 51\tnote{1} & 60 & 63 \\
  & PD\_Onset & 55.9 (10.8) & 56.2 (10.4) & 56.7 (10.5) & 57.8 (11.0) \\
  & PD\_Duration & 6.8 (4.8)\tnote{34} & 8.3 (5.3)\tnote{4} & 9.3 (6.3)\tnote{14} & 13.8
  (8.2)\tnote{123} \\
  & CISI\_Total & 6.3 (3.4)\tnote{234} & 10.8 (4.7)\tnote{14} & 10 (4.0)\tnote{14} & 15.1 (5.7)\tnote{123} \\
   \bottomrule
\end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{nms30m-colhc-pub.pdf}
  \caption{Hierarchical clustering of motor (blue) and nonmotor (black) symptoms. Symptoms are
  labeled with their name and corresponding domain number. Dendrogram colored with 4 clusters.}
  \label{fig:hc}
\end{figure*}

% TODO: If the outline of the paper is good enough it might be possible to have this inline
% in the paper somewhere. If so, change figure* -> figure, make h! (force horizontal?), and switch
% 0.8 linewidth to full linewidth
% Also, I'm going to include this AFTER nms30m-colhc, so that colhc/boxplots fits nicely on one
% page
\begin{figure*}[p]
  \centering
  \includegraphics[width=0.7\linewidth]{nms30-hm-pub.pdf}
  % TODO: Variables not used in analysis? Label color - black nms, blue nonmotor, grey ?
  \caption{Heatmap of symptoms clustering. Color indicates cluster mean as a $z$-score for the
  given symptom.}
  \label{fig:nms30-hm}
\end{figure*}

\subsection{Hierarchical clustering on symptoms}

Hierarchical clustering on the 30 nonmotor symptoms and the four motor symptoms produced the
dendrogram in Figure~\ref{fig:hc}. Predictably, symptoms in the same nonmotor domains tended to
cluster together, with the exceptions of diplopia (domain 4), drowsiness (domain 2), and the
symptoms in domain 9 (miscellaneous). Notably, tremor was the most dissimilar symptom, occupying a
single branch at the top of the tree.

\subsection{Longitudinal analysis}

Most variables had little to no correlation with PD duration (Figure~\ref{fig:corr}). Variables
with the highest correlation include symptoms urinary frequency, swallowing, and drooling; domains
miscellaneous, urinary, and gastrointestinal; and total CISI score. Scatterplots for CISI Total,
Tremor, Anxiety, and Depression appear in Figure~\ref{fig:long4}.

\begin{figure*}[p]
  \centering
  \includegraphics[width=\linewidth]{cor-unbinned.pdf}
  \caption{Linear correlation of each symptom with PD duration.}
  \label{fig:corr}
\end{figure*}

\begin{figure*}[p]
  \centering
  \includegraphics[width=0.8\linewidth]{long4-d.pdf}
  \caption{Selected symptoms for 901 patients plotted against PD duration. Smoothed loess curves for each cluster
  are drawn in their respective colors; the black curve is the curve for the entire population.}
  \label{fig:long4}
\end{figure*}

\section{Discussion}

Domains clustering reveals clusters that confirm previous
findings in the field, mainly van Rooden et al.\cite{vanrooden10} and the identification of four
subtypes of Parkinson's disease: mild, nonmotor-dominant, motor-dominant, and severe. van
Rooden's work was conducted with a separate dataset using a different modeling method
(expectation-maximization), and this investigation independently confirms these subtype
classifications. Unlike van Rooden, mean disease durations differences do exist between subtypes 1
(mild) and 4 (severe), likely due to further development of the disease, although the differences
between 2 and 3 (nonmotor/motor predominated) subtypes are insignificant
(Table~\ref{tab:nmd}), importantly suggesting different developmental paths of the disease.

Overall, little information was found in pdonset,
durat\_pd, or current age. Mean ages were similar for clusters 1, 2, and 3 ($p >
0.05$), but different for the severe cluster 4, intuitively since cluster 4 represents a more
advanced stage of the disease.  Specifically, clusters 1 and 4 seem to be phenotypically quite
similar, except at different stages of disease progression, given cluster 4's higher age and
durat\_pd scores.

However, clusters 2 and 3 clearly show different disease progression, one in
the motor direction, and one in the nonmotor. Both groups have similar age,
PD onset, and PD duration scores, but differ significantly in symptomatic expression.
Cluster 2 is dominated by a high severity of nonmotor domains, especially Sleep/Fatigue,
Mood/Cognition, Urinary, and Miscellaneous.  Cluster 3, however, is dominated by a high prevalence
of motor symptoms, where most motor symptoms are similar to the mild cluster 1. Of note is that the
tremor population mean is the highest cluster mean, even higher than the severe subtype 4. This
motor-dominant cluster may thus overlap with Ma's tremor dominant/slow progression
cluster\cite{ma15}.

Generally, given stable pdonset scores and predictably increasing durat\_pd
scores for clusters 1 and 4, Ma et al's rapid disease
progression/late onset and tremor dominant/slow progression clusters
\cite{ma15} were mostly not found in this dataset, save for the tremor-dominant
motor cluster.

% The most important nonmotor symptoms in determining these clusters were nms\_d2
% (sleep) and nms\_d3 (mood/cognition), which echo findings of Fereshtehnejad's
% longitudinal study\cite{fereshtehnejad15} and are similar to Sauerbier's
% identification of sleep dominant and cognitive dominant clinical NMS subtypes\cite{sauerbier15}.
% Compared to Erro et al.\cite{erro13}, nonmotor/motor
% dominant subtypes were indeed found, but an additional subgroup with relatively
% severe levels of both motor and nonmotor symptoms were found. Erro's benign
% subtype groups possibly overlap with the mild cluster 1 found in this
% investigation.

% \subsection{Nonmotor subtype: clustering and modeling}

% Nonmotor symptoms nms\_d2 and nms\_d3 became critical not only in
% classification trees distinguishing between the various symptoms but in the
% nonmotor-predominant subgroup itself. In $k$-means subdivision of the
% nonmotor-dominant subtype where $k = 2$ and $k = 3$, opposite trends were
% confirmed with nms\_d2 and nms\_d3 symptoms. Similarly, in the 2 and 4 vs rest
% decision tree (Figure~\ref{fig:dtree-2and4va-pruned}), nms\_d2 and nms\_d3 nodes were
% used to differentiate various categories of nomotor-dominant patients.
% When $k = 3$, the subtype with the highest nms\_d2 scores and lowest
% nms\_d3 scores had by far the highest axial scores, nms\_d6 (gastrointestinal)
% scores, and nms\_d7 (urinary) scores. Thus subtype 3 of the nonmotor-dominated
% group could include patients falling into the cognitive/depression-dominant or
% autonomic dominant subtypes.

% Despite the variety in symptomatic expression in this nonmotor group, what
% seems most consistent is the presence of nms\_d9 (miscellaneous) nonmotor
% symptoms, as it is used as the root node of the 2 vs all decision tree
% (Figure~\ref{fig:2va}) and the 2 and 4 vs rest decision tree
% (Figure~\ref{fig:dtree-2and4va-pruned}).

Our longitudinal analysis gives more insight into the clusters found in the previous analyses.
According to Figure~\ref{fig:corr}, most symptoms are uncorrelated with PD duration, especially
mood/cognition symptoms (anxiety, depression) and tremor.
The differences in disease progression for each cluster can be seen by the corresponding graphs in
Figure~\ref{fig:long4}. Cluster 2 (Nonmotor-Dominant) starts at higher
scores for anxiety and depression, and actually decreases with increasing PD duration, thus indicating that these patients' subtype can be determined
early after PD onset from the depressive symptom score. Similarly, when examining cluster 3
(Motor-dominant), the mean tremor score is substantially higher from PD onset.
Interestingly, cluster 4 (Severe) generally starts at lower tremor and motor scores during disease,
but then rises sharply, exceeding other clusters. More evidence that tremor is a unique motor
symptom is located in Figure~\ref{fig:hc}, where it is the most distant symptom from all other
symptoms.

% Indeed, PCA on the 30 nonmotor symptoms identifies the second-most prominent component as a general
% mood/cognition component, and $k$-means clustering on the 30 symptoms only
% (Figure~\ref{fig:nms30m-k4-heatmap-improved}) divides the 1000 patients into four slightly
% different groups, a mild, average, depression-dominant, and severe group.

% The Gaussian mixture model identified in Figure~\ref{fig:nms30m-mclust-heatmap-improved} fragments
% the previously-discovered clusters into more groups. Here, a wide variety of specialized subtypes
% of PD are displayed, including insomnia, urinary, motor, nonmotor, and depression-dominant groups,
% as well as the expected mild, average, and severe subtypes. It is likely that the previous analysis
% with only nms\_d\{1-9\} combined most of those specialized groups into the more general
% nonmotor-dominant Subtype 2.

It's intuitive that a Depression-Dominant group emerges when clustering on nms\{1-30\}, since
domain 3 consists of 5 separate measures. Thus, any high expression of depressive symptoms is
magnified in clustering, since the symptoms are highly similar (Figure~\ref{fig:hc}) and
treated with equal weight. Once again reinforcing what was discovered previously, depressive
symptoms have been shown to be very important in determining subtypes of PD.

\section{Supplementary Material}

Probably the large table goes here?

\section{References}

\bibliography{manuscript}

\end{document}